Reproducible Research Project 1
========================================================
```{r}
setwd('C:/Users/Ritesh/Desktop/projects/R Working Dir/Reproducible Research/Project 1')
#Show any code that is needed to
#1.Load the data (i.e. read.csv())

    activity=read.csv('activity.csv')
    #1 Daily Activity
      #.Process/transform the data (if necessary) into a format suitable for your analysis
      #What is mean total number of steps taken per day?
      #For this part of the assignment, you can ignore the missing values in the dataset.
      #1.Make a histogram of the total number of steps taken each day
        y=aggregate(activity$steps~activity$date,FUN="sum")
        x11()
        colnames(y)=c("Date","Steps")
        hist(y$Steps,xlab="# Steps taken",ylab="# Days",main="Histogram of Number of Days grouped by Steps taken")
    #2.Calculate and report the mean and median total number of steps taken per day
      #Mean= 10766.19, Median=10765, code is below
        Mean=mean(y$Steps) 
        Median=median(y$Steps) 
      #What is the average daily activity pattern?
      #1.Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
        x=aggregate(steps~interval,data=activity,FUN="mean")
        plot(x$interval,x$steps,type='l',xlab="Interval",ylab="Steps",main="#Steps by interval")
      #2.Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
        #104th interval corresponding to 835th minute
        x$interval[which(x$steps==max(x$steps),x$interval)]
```
#2 Imputing missing values
    #Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.
    #1.Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
      # Total 2304 missing values
```{r}
      nrow(activity[activity$steps=='NA',])
    #2.Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
      #Replaced the NAs with Average for that interval
      nastep=activity[which(is.na(activity$step),TRUE),]
      nonna=activity[which(!(is.na(activity$step)),TRUE),]
      nasteplkup=merge(nastep,x,by="interval")
      lkupfinal=nasteplkup[,c("steps.y","date","interval")]
      colnames(lkupfinal)=c("steps","date","interval")
    #3.Create a new dataset that is equal to the original dataset but with the missing data filled in.
      merged=rbind(lkupfinal,nonna)
    #4.Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
      mergedgrouped=aggregate(merged$steps~merged$date,FUN="sum")
      colnames(mergedgrouped)=c("Date","Steps")
      x11()
      hist(mergedgrouped$Steps,xlab="# Steps taken",ylab="# Days",main="Histogram of Number of Days grouped by Steps taken")
      #Do these values differ from the estimates from the first part of the assignment? 
      #Mean remains the same while median has changed.
      Mean1=mean(mergedgrouped$Steps)
      Median1=median(mergedgrouped$Steps)
      #What is the impact of imputing missing data on the estimates of the total daily number of steps?
      #Total steps per day has increased
      sum(mergedgrouped$Steps,na.rm=TRUE)>sum(activity$steps,na.rm=TRUE)
```
#3 Are there differences in activity patterns between weekdays and weekends?
    #For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.
    #1.Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
    merged$date=as.Date(merged$date)
    mergedfinal=cbind(weekdays(merged$date),merged)
    library(timeDate)
    library(ggplot2)
    temp=isWeekday(merged$date)
    temp=gsub("TRUE","Weekday",temp)
    temp=gsub("False","Weekend",temp)
    mergedfinal1=cbind(mergedfinal,temp)
    colnames(mergedfinal1)[5]="IsWkday"
    groupedmerged=aggregate(mergedfinal1$steps~mergedfinal1$IsWkday+mergedfinal1$interval,FUN="mean")
    #2.Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). The plot should look something like the following, which was creating using simulated data:
```   
    x11()
    colnames(groupedmerged)=c("IsWkday","interval","steps")
    qplot(x=interval,y=steps,data=groupedmerged,facets=IsWkday~.,geom="line")
      
   
